---
title: "EPA Data Download v5"
author: "Jeffrey Wu"
date: "2023-09-02"
output: pdf_document
---

LOAD LIBRARIES: 

```{r}
library(tidyverse)
library(plyr)
library(dplyr)
library(jsonlite)
library(lubridate)
library(ggplot2)
library(maps)
library(mapdata)
library(geosphere)
library(urbnmapr)
library(RAQSAPI)
library(con2aqi)

aqs_credentials("jeffreywu@ucsb.edu","goldswift19")
```

GET COUNTY AND POLLUTANT REFERENCE CODES FROM EPA

```{r}
#Get county codes 
counties_url = "https://aqs.epa.gov/data/api/list/countiesByState?email=jeffreywu@ucsb.edu&key=goldswift19&state=06&"

countycodes = fromJSON(counties_url)
countycodes = countycodes[[2]]
california_counties = countycodes$code

#Get parameter codes
parameters_url = "https://aqs.epa.gov/data/api/list/parametersByClass?email=jeffreywu@ucsb.edu&key=goldswift19&pc=CRITERIA"

parametercodes = fromJSON(parameters_url)
parametercodes = parametercodes[[2]]

#Do we want to keep Lead PM10 or delete
parametercodes = parametercodes[-7,]

pollutants = data.frame(parametercodes$code)
labels = c("lead","co","so2","no2","o3","pm10","pm25")
pollutants = cbind(pollutants,labels)
```


##IDENTIFYING SET OF MONITORING STATIONS THAT ARE WELL DISTRIBUTED ACROSS CALIFORNIA:

FUNCTION THAT QUERIES STATION LOCATIONS FOR A GIVEN POLLUTANT 

```{r}
query_aqs_station_data <- function(param,year){
  start_date <- paste0(year, "0101")
  end_date <- paste0(year, "1231")
  
  url <- paste0("https://aqs.epa.gov/data/api/monitors/byState?email=jeffreywu@ucsb.edu&key=goldswift19&param=", param, "&bdate=", start_date, "&edate=", end_date, "&state=06")
  
  myData <- fromJSON(url)
  station_data = myData[[2]]
  
  return(station_data)
}
```


FOR EACH PARAMETER, GRAB ALL MONITORING STATIONS FOR EACH YEAR

```{r}
stations_url = "https://aqs.epa.gov/data/api/monitors/byState?email=jeffreywu@ucsb.edu&key=goldswift19&param=88101&bdate=20140101&edate=20141231&state=06"

stations = fromJSON(stations_url)
stations = stations[[2]]

station_data2014_pm2.5 = stations %>% select(latitude,longitude,site_number, local_site_name,county_code,county_name)

#Get monitoring station locations for each pollutant for each year (takes approx 3 min)
all_pollutants_station_data <- list()
for (year in 2014:2019){
  year_data <- lapply(pollutants, query_aqs_station_data, year = year)
  all_pollutants_station_data[[as.character(year)]] <- year_data
}


# 1 - Lead, 2 - Carbon monoxide (CO), 3 - Sulfure dioxide (SO2), 4 - Nitrogen dioxide (NO2)
# 5 - Ozone (O3), 6 - Total PM10, 7 - PM2.5

all_pollutants_station_2014data = all_pollutants_station_data[[1]]
all_pollutants_station_2015data = all_pollutants_station_data[[2]]
all_pollutants_station_2016data = all_pollutants_station_data[[3]]
all_pollutants_station_2017data = all_pollutants_station_data[[4]]
all_pollutants_station_2018data = all_pollutants_station_data[[5]]
all_pollutants_station_2019data = all_pollutants_station_data[[6]]
```


FUNCTION TO PLOT MONITORING STATIONS FOR A GIVEN POLLUTANT ON A MAP OF CA

```{r}
# Function to plot latitude and longitude on a map of California with county lines
plot_latitude_longitudes <- function(latitude, longitude) {
  # Get the county map data for California
  county_map <- map_data("county", region = "california")
  
  # Create a data frame with latitude and longitude
  data <- data.frame(lat = latitude, lon = longitude)
  
  # Plot the map with county lines and points
  p <- ggplot() +
    geom_polygon(data = county_map, aes(x = long, y = lat, group = group), color = "black", fill = "lightgray") +
    geom_point(data = data, aes(x = lon, y = lat), color = "red", size = 2) +
    scale_fill_discrete()
    labs(title = "Latitude and Longitude in California with County Boundaries") +
    theme_minimal()
  
  print(p)
}

# Define latitude and longitude vectors
# simple test 
# latitude <- c(36.7783, 34.0522, 33.6846)
# longitude <- c(-119.4179, -118.2437, -117.8265) 

#Plot monitoring station locations for a specific pollutant and specific year
# 1 - Lead, 2 - Carbon monoxide (CO), 3 - Sulfure dioxide (SO2), 4 - Nitrogen dioxide (NO2)
# 5 - Ozone (O3), 6 - Total PM10, 7 - PM2.5

latitude = unique(all_pollutants_station_2014data[[3]]$latitude)
longitude = unique(all_pollutants_station_2014data[[3]]$longitude)
CA = plot_latitude_longitudes(latitude, longitude)

plot_latitude_longitudes(test$latitude, test$longitude)
```


IDENTIFY SUBSET OF STATIONS THAT HAS BEST SPATIAL COVERAGE WRT CENTROID OF CA STATE 

Lat/long for 2 biggest cities (based on population) in each county 

```{r}
citylats = c(37.8044,37.5485,38.7743,38.8071,38.3527,38.3488,39.7285,39.7596,
             38.1231,38.0678,39.2143,39.1546,37.9780,38.0049,41.7558,41.7548,
             38.9399,38.6688,36.7378,36.8252,39.7474,39.5243,40.8021,40.8665,
             32.7920,32.6789,37.3614,37.3855,35.3733,35.7688,36.3275,36.3008,
             38.9582,38.8080,40.4163,40.2840,34.0522,33.7701,36.9613,37.1230,
             37.9735,38.1074,37.4849,37.4320,39.4457,39.4096,37.3022,37.0583,
             41.4871,41.4099,37.6485,38.5149,36.6777,36.6149,38.2975,38.1749,
             39.3280,39.2191,33.8366,33.7455,38.7521,38.7907,39.9341,40.3063,
             33.9806,33.9425,38.5816,38.4088,36.8525,36.8125,34.1083,34.0922,
             32.7157,32.6401,37.7749,37.9780,37.9577,37.7396,35.2828,35.6369,
             37.6879,37.5630,34.9530,34.4208,37.3387,37.3688,36.9741,36.9102,
             40.5865,40.4482,39.6763,39.5595,41.7354,41.3099,38.1041,38.2492,
             38.4404,38.2324,37.6393,37.4946,39.1404,39.1165,40.1785,39.9277,
             40.7310,40.4156,36.3301,36.2077,38.0297,37.9829,34.1975,34.1706,
             38.5449,38.6785,39.1277,39.0954)

citylongs = c(122.2712,121.9886,119.8219,119.7960,120.9327,120.7741,121.8375,121.6219,
              120.8509,120.5385,122.0094,122.1494,122.0311,121.8058,124.2026,124.1580,
              119.9772,120.9872,119.7871,119.7029,122.1964,122.1936,124.1637,124.0828,
              115.5631,115.4989,118.3997,118.4105,119.0187,119.2471,119.6457,119.7829,
              122.6264,122.5583,120.6530,120.5394,118.2437,118.1937,120.0607,120.2602,
              122.5311,122.5697,119.9663,120.0985,123.8053,123.3556,120.4830,120.8499,
              120.5425,120.6791,118.9721,119.4768,121.6555,121.8221,122.2869,122.2608,
              120.1833,121.0611,117.9143,117.8677,121.2880,121.2358,120.8980,121.2319,
              117.3755,117.2297,121.4944,121.3716,121.4016,121.3658,117.2898,117.4350,
              117.1611,117.0842,122.4194,122.0311,121.2908,121.4260,120.6596,120.6545,
              122.4702,122.3255,120.4357,119.6982,121.8853,122.0363,122.0308,121.7569,
              122.3917,122.2978,120.2410,120.8277,122.6345,122.3106,122.2566,122.0405,
              122.7141,122.6367,120.9970,120.8460,121.6169,121.6380,122.2358,122.1792,
              122.9420,123.2100,119.2966,119.3473,119.9741,120.3822,119.1771,118.8376,
              121.7405,121.7733,121.5508,121.5522)
citylongs = -1*citylongs
```


Alameda: Oakland (429082) and Fremont 

Alpine: Alpine Village (225) and Mesa Vista 

Amador: Ione (8363) and Jackson

Butte: Chico (94776) and Paradise 

Calaveras: Rancho Calaveras (5324) and Angels Camp

Colusa: Colusa (5911) and Williams 

Contra Costa: Concord (129688) and Antioch 

Del Norte: Crescent City (6805) and Bertsch-Oceanview 

El Dorado: South Lake Tahoe (22036) and Cameron Park 

Fresno: Fresno (530093) and Clovis 

Glenn: Orland (7644) and Willows 

Humboldt: Eureka (26998) and Arcata

Imperial: El Centro (44120) and Calexico 

Inyo: Bishop (3746) and Dixon Lane-Meadow Creek

Kern: Bakersfield (383579) and Delano 

Kings: Hanford (56910) and Lemoore

Lake: Clearlake (15384) and Hidden Valley Lake 

Lassen: Susanville (15165) and Janesville

Los Angeles: Los Angeles (3990000) and Long Beach 

Madera: Madera (65706) and Chowchilla

Marin: San Rafael (58704) and Novato 

Mariposa: Mariposa (1526) and Catheys Valley

Mendocino: Fort Bragg (7359) and Willits 

Merced: Merced (83316) and Los Banos 

Modoc: Alturas (2509) and California Pines

Mono: Mammoth Lakes (8127) and Walker

Monterey: Salinas (156259) and Seaside

Napa: Napa (79263) and American Canyon 

Nevada: Truckee (16561) and Grass Valley 

Orange: Anaheim (352005) and Santa Ana 

Placer: Roseville (139117) and Rocklin

Plumas: East Quincy (2489) and Chester 

Riverside: Riverside (330063) and Moreno Valley 

Sacramento: Sacramento (508529) and Elk Grove

San Benito: Hollister (39749) and Ridgemark 

San Bernandino: San Bernandino (215941) and Fontana

San Diego: San Diego (1426000) and Chula Vista 

San Francisco: San Francisco (810000) and Concord 

San Joaquin: Stockton (311178) and Tracy

San Luis Obispo: San Luis Obispo (47446) and Paso Robles 

San Mateo: Daly City (107008) and San Mateo 

Santa Barbara: Santa Maria (107408) and Santa Barbara 

Santa Clara: San Jose (1030000) and Sunnyvale 

Santa Cruz: Santa Cruz (64725) and Watsonville 

Shasta: Redding (91772) and Anderson

Sierra: Loyalton (700) and Downieville 

Siskiyou: Yreka (7556) and Mount Shasta

Solano: Vallejo (121913) and Fairfield 

Sonoma: Santa Rosa (177586) and Petaluma

Stanislaus: Modesto (215030) and Turlock

Sutter: Yuba City and South Yuba City 

Tehama: Red Bluff (14283) and Corning 

Trinity: Weaverville (3667) and Post Mountain

Tulare: Visalia (133800) and Tulare 

Tuolumne: Phoenix Lake-Cedar Ridge (5108) and Sonora 

Ventura: Oxnard (209877) and Thousand Oaks 

Yolo: Davis (69289) and Woodland

Yuba: Linda (17773) and Olivehurst



FUNCTION THAT SELECTS SET OF STATIONS CLOSEST TO A GIVEN LAT/LONG 

```{r}
# Function to filter stations based on spatial coverage
subset_stations_by_spatial_coverage <- function(station_data, reference_lat, reference_lon, max_distance_km=100) {
  # Calculate distances between stations and reference location
  distances <- distHaversine(
    cbind(station_data$longitude, station_data$latitude),
    c(reference_lon, reference_lat)
  )
  distances <- distances/1000
  
  # idx =  which(distances == min(distances))
  # #Identify station within min distance to centroid of county
  # station_data_subset <- station_data[idx, ]
  
  # Subset stations within the specified max_distance_km
  idx = which(distances <= max_distance_km)
  station_data_subset <- station_data[idx, ]
  station_data_subset <- cbind(station_data_subset,distances[idx])
  
  while (nrow(station_data_subset) < 5){
    max_distance_km = max_distance_km + 50
    station_data_subset = subset_stations_by_spatial_coverage(station_data, 
                              reference_lat, reference_lon, max_distance_km)
  }
  
  return(station_data_subset)
}

# Construct the subset of stations based on spatial coverage criteria (test)
reference_lat = citylats[3]
reference_lon = citylongs[3]
max_distance_km = 100

subset_stations <- subset_stations_by_spatial_coverage(station_data2014_pm2.5, reference_lat, reference_lon, max_distance_km)

# Print the subset of stations
print(subset_stations)
```


```{r}
# Obtain centroid lat/longs for each county 
CA.counties2 = read.csv("counties.ca.data.csv")
ca.coordinates = data.frame(CA.counties2$county,CA.counties2$lat,CA.counties2$lng)
colnames(ca.coordinates) = c("county","lat","long")

ca.coordinates = ca.coordinates[order(ca.coordinates$county),]
row.names(ca.coordinates) = NULL
```


GIVEN A DATASET CONTAINING STATION LOCATIONS FOR A GIVEN YEAR, THIS FUNCTION SELECTS A SET OF STATIONS (5-20) THAT ARE CLOSEST TO THE LAT/LONGS FOR THE TWO BIGGEST CITIES IN EACH COUNTY FOR A GIVEN POLLUTANT AND MAKES A DATAFRAME OUT OF IT 

```{r}
#Function that finds best monitoring station for each county for a specific pollutant for a specific year
# 1 - Lead, 2 - Carbon monoxide (CO), 3 - Sulfure dioxide (SO2), 4 - Nitrogen dioxide (NO2)
# 5 - Ozone (O3), 6 - Total PM10, 7 - PM2.5

best_stations = function(stationdata,pollutant){

  subset_list = list()
  
  #Load lat/longs for 58x2 cities into dataframe
  CA.coords = data.frame(rep(countycodes$value_represented,each = 2),citylats,citylongs)
  colnames(CA.coords) = c("County","Lat","Long")
  
  #Find closest station for each county centroid using subset_stations_by_spatial_coverage function
  for (i in 1:nrow(CA.coords)){
  reference_lat = CA.coords$Lat[i]
  reference_lon = CA.coords$Long[i]
  max_distance_km = 100
  
  subset_stations <- subset_stations_by_spatial_coverage(stationdata[[pollutant]], reference_lat, reference_lon, max_distance_km)
  subset_list[[i]] = subset_stations
  }
  
  #Combine pairs of city lists together 
  subset_list2 = list()
  sequence = seq(2,116,2)
  for(i in sequence){
    combine = rbind(subset_list[[i]],subset_list[[i-1]])
    subset_list2[[i-1]] = combine
  }
  subset_list2 =subset_list2[!sapply(subset_list2,is.null)]
  
  #Create a county label vector
  repnames = c()
  for(i in 1:58){
    repnames = c(repnames,nrow(subset_list2[[i]]))
  }  
  countylabels = rep(countycodes$value_represented,times = repnames)
  
  #Format the list into dataframe
  beststations = as.data.frame(do.call(rbind, subset_list2))
  beststations = cbind(countylabels,beststations$county_name,
                       beststations$`distances[idx]`,beststations)
  colnames(beststations)[c(1,2,3)] = c("measuring_county","station_county","distance_apart")
  rownames(beststations) = NULL
  
  return(beststations)
}

#test cases
pm2.5_stations_2014 = best_stations(all_pollutants_station_2014data,7)
CO_stations_2016 = best_stations(all_pollutants_station_2016data,2)
```


CREATING BEST STATION LIST/DATAFRAME FOR EACH POLLUTANT, EACH ENTRY IS A YEAR

```{r}
#Generate list for best stations for each pollutant for each year
Lead_stations = list()

Lead_stations[[1]] = best_stations(all_pollutants_station_2014data,1)
Lead_stations[[2]] = best_stations(all_pollutants_station_2015data,1)
Lead_stations[[3]] = best_stations(all_pollutants_station_2016data,1)
Lead_stations[[4]] = best_stations(all_pollutants_station_2017data,1)
Lead_stations[[5]] = best_stations(all_pollutants_station_2018data,1)
Lead_stations[[6]] = best_stations(all_pollutants_station_2019data,1)



CO_stations = list()

CO_stations[[1]] = best_stations(all_pollutants_station_2014data,2)
CO_stations[[2]] = best_stations(all_pollutants_station_2015data,2)
CO_stations[[3]] = best_stations(all_pollutants_station_2016data,2)
CO_stations[[4]] = best_stations(all_pollutants_station_2017data,2)
CO_stations[[5]] = best_stations(all_pollutants_station_2018data,2)
CO_stations[[6]] = best_stations(all_pollutants_station_2019data,2)



SO2_stations = list()

SO2_stations[[1]] = best_stations(all_pollutants_station_2014data,3)
SO2_stations[[2]] = best_stations(all_pollutants_station_2015data,3)
SO2_stations[[3]] = best_stations(all_pollutants_station_2016data,3)
SO2_stations[[4]] = best_stations(all_pollutants_station_2017data,3)
SO2_stations[[5]] = best_stations(all_pollutants_station_2018data,3)
SO2_stations[[6]] = best_stations(all_pollutants_station_2019data,3)



NO2_stations = list()

NO2_stations[[1]] = best_stations(all_pollutants_station_2014data,4)
NO2_stations[[2]] = best_stations(all_pollutants_station_2015data,4)
NO2_stations[[3]] = best_stations(all_pollutants_station_2016data,4)
NO2_stations[[4]] = best_stations(all_pollutants_station_2017data,4)
NO2_stations[[5]] = best_stations(all_pollutants_station_2018data,4)
NO2_stations[[6]] = best_stations(all_pollutants_station_2019data,4)



O3_stations = list()

O3_stations[[1]] = best_stations(all_pollutants_station_2014data,5)
O3_stations[[2]] = best_stations(all_pollutants_station_2015data,5)
O3_stations[[3]] = best_stations(all_pollutants_station_2016data,5)
O3_stations[[4]] = best_stations(all_pollutants_station_2017data,5)
O3_stations[[5]] = best_stations(all_pollutants_station_2018data,5)
O3_stations[[6]] = best_stations(all_pollutants_station_2019data,5)



PM10_stations = list()

PM10_stations[[1]] = best_stations(all_pollutants_station_2014data,6)
PM10_stations[[2]] = best_stations(all_pollutants_station_2015data,6)
PM10_stations[[3]] = best_stations(all_pollutants_station_2016data,6)
PM10_stations[[4]] = best_stations(all_pollutants_station_2017data,6)
PM10_stations[[5]] = best_stations(all_pollutants_station_2018data,6)
PM10_stations[[6]] = best_stations(all_pollutants_station_2019data,6)



# Lead.PM10_stations = list()
# 
# Lead.PM10_stations[[1]] = best_stations(all_pollutants_station_2014data,7)
# Lead.PM10_stations[[2]] = best_stations(all_pollutants_station_2015data,7)
# Lead.PM10_stations[[3]] = best_stations(all_pollutants_station_2016data,7)
# Lead.PM10_stations[[4]] = best_stations(all_pollutants_station_2017data,7)
# Lead.PM10_stations[[5]] = best_stations(all_pollutants_station_2018data,7)
# Lead.PM10_stations[[6]] = best_stations(all_pollutants_station_2019data,7)



PM2.5_stations = list()

PM2.5_stations[[1]] = best_stations(all_pollutants_station_2014data,7)
PM2.5_stations[[2]] = best_stations(all_pollutants_station_2015data,7)
PM2.5_stations[[3]] = best_stations(all_pollutants_station_2016data,7)
PM2.5_stations[[4]] = best_stations(all_pollutants_station_2017data,7)
PM2.5_stations[[5]] = best_stations(all_pollutants_station_2018data,7)
PM2.5_stations[[6]] = best_stations(all_pollutants_station_2019data,7)
```



##DOWNLOADING AND AGGREGATING AIR QUALITY DATA USING DIRECT API CALLS

CREATE A SET OF 5-20 MONITORING STATIONS FOR EACH COUNTY AND LOOP THROUGH THOSE STATION CODES 

QUERYING MULTIPLE PARAMETERS AT ONCE: 

```{r}
# 1 - Lead, 2 - Carbon monoxide (CO), 3 - Sulfure dioxide (SO2), 4 - Nitrogen dioxide (NO2)
# 5 - Ozone (O3), 6 - Total PM10, 7 - PM2.5

stations2019x = rbind(Lead_stations[[6]],CO_stations[[6]],
             SO2_stations[[6]],NO2_stations[[6]])
stations2019y = rbind(O3_stations[[6]],PM10_stations[[6]],PM2.5_stations[[6]])

sitenums2019x = stations2019x %>% select(county_code,site_number) %>% unique() #198 stations
sitenums2019y = stations2019y %>% select(county_code,site_number) %>% unique() #178 stations


#Trying EPA R Package query (took 15 + 20 min!) gives us a dataframe 
ccodes = sitenums2019y$county_code
snums = sitenums2019y$site_number
str1 = "2019-01-01"
str2 = "2019-12-31"

pollutants1_2019 = aqs_dailysummary_by_site(parameter = c("14129","42101","42401","42602"),bdate = as.Date(str1),edate = as.Date(str2),stateFIPS = "06",countycode = ccodes,sitenum = snums)

pollutants2_2019 = aqs_dailysummary_by_site(parameter = c("44201","81102","88101"),bdate = as.Date(str1),edate = as.Date(str2),stateFIPS = "06",countycode = ccodes,sitenum = snums)

###SAVE LIST LOCALLY
saveRDS(pollutants2_2019,file = "C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Raw/pollutants2_2019_8.18.RData")
```

QUALITY CHECK FUNCTION FOR STATION DATA: WANT TO ADDRESS OUTLIERS, MISSINGNESS 

```{r}
#Given a dataset like CO2016 (list of 1000ish stations), check for 2/3 missing data and for outliers

station_quality_check = function(station_data){
  l = length(station_data)
  badindex = c()
  consecutive_outliers = list()
  
  for (i in 1:l){
    aqi = station_data[[i]]$aqi
    pollutant_level = station_data[[i]]$arithmetic_mean
    
    #check for outliers in AQI
    median.aqi = median(na.omit(aqi))
    mad.aqi = mad(na.omit(aqi))
    
    min.aqi = median.aqi-(3*mad.aqi)
    max.aqi = median.aqi+(3*mad.aqi)
    
    outliers.aqi = which(aqi < min.aqi | aqi > max.aqi)
    
    result.aqi = rle(diff(outliers.aqi))

    
    #check for outliers in pollutant measure
    median.pollutant = median(na.omit(pollutant_level))
    mad.pollutant = mad(na.omit(pollutant_level))
    
    min.pollutant = median.pollutant-(3*mad.pollutant)
    max.pollutant = median.pollutant+(3*mad.pollutant)
    
    outliers.pollutant = which(pollutant_level < min.pollutant | pollutant_level > max.pollutant)
    
    result.pollutants = rle(diff(outliers.pollutant))
    
    
    if (nrow(station_data[[i]]) < 240){
      badindex = c(badindex,i)
    }
    
    else if (any(result.aqi$lengths >= 14 & result.aqi$values == 1) == TRUE){
      badindex = c(badindex,i)
    }
    
    else if (any(result.pollutants$lengths >= 14 & result.pollutants$values == 1) == TRUE){
      badindex = c(badindex,i)
    }
    
    consecutive_outliers[[i]] = c("AQI",outliers.aqi,"POLLUTANTS",outliers.pollutant)
  }
  
  bad_list = list(badindex,consecutive_outliers)
  
  return(bad_list)
}


#Test on CO2016 and CO2017
# station_quality_check(CO2016) #returns 61 "bad stations" out of 1230
# station_quality_check(CO2014) #returns 495 "bad stations" out of 1230 (old approach)
# 
# removeidx = station_quality_check(SO22017)[[1]] #returns 715 out of 1056 "bad stations" 
# 
# test = SO22017[- removeidx]
```

The station check for SO22017 removed all the stations for SB, can't have that... can we? Or maybe it's ok because it just means SB doesn't have any replications for SO2 in 2017? 


FUNCTION THAT AGGREGATES DAILY DATA INTO MONTHYLY MEDIANS 

```{r}
monthly_agg = function(pollutantdata){
  #Aggregating all the station data at once
  date = ymd(pollutantdata$date_local)
  df2 <- pollutantdata                                   # Duplicate data
  df2$year_month <- floor_date(date,"month")  # Create year-month column
  df3 = df2 %>% select(county,site_number,arithmetic_mean,aqi,year_month) %>% as.data.frame()
  
  df3$arithmetic_mean = as.numeric(df3$arithmetic_mean)
  df3$aqi[which(df3$aqi == "NULL")] = NA
  df3$aqi = as.numeric(df3$aqi)
  
  df.agg = df3 %>% group_by(year_month) %>% dplyr::summarize(arithmetic_mean = median(na.omit(arithmetic_mean)),aqi = median(na.omit(aqi))) %>% as.data.frame()
  
  return(df.agg)
}
```


END GOAL IS TO GET TIME SERIES OF DATA FOR EACH POLLUTANT
CURRENT PROPOSED FINAL FORM: ONE BIG DATAFRAME (ALL POLLUTANTS ALL YEARS TOGETHER, USE FILTER TO SEPARATE)


# Function transforming raw data to final form 

```{r}
# Group 1: 14129 - Lead, 421012 - Carbon monoxide (CO), 42401 - Sulfure dioxide (SO2), 42602 - Nitrogen dioxide (NO2)
# Group 2: 44201 - Ozone (O3), 81102 - Total PM10, 88101 - PM2.5

pollutants1_2014 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Raw/pollutants1_2014_8.17.RData")
pollutants2_2014 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Raw/pollutants2_2014_8.17.RData")

CO2014 = pollutants1_2014 %>% filter(parameter_code == "42101")

raw_transform = function(rawdata,reference_list,standard){
  
  ###SEPARATE DF INTO A LIST OF DFs 
  
  matched_list = list()
  
  if(missing(standard)){
    for (i in 1:nrow(reference_list)){
    data = rawdata %>% filter(county_code == reference_list$county_code[i], site_number == reference_list$site_number[i])
  
    matched_list[[i]] = data
    }
  } else {
      for (i in 1:nrow(reference_list)){
      data = rawdata %>% filter(county_code == reference_list$county_code[i], site_number == reference_list$site_number[i],pollutant_standard == standard)

      matched_list[[i]] = data
    }
  }
  
  names(matched_list) = reference_list$measuring_county
  
  ###STATION QUALITY CHECK
  
  removeidx = station_quality_check(matched_list)[[1]]
  good_matched_list = matched_list[- removeidx]
  
  #Convert list back into one big dataframe
  temp = as.data.frame(do.call(rbind, good_matched_list)) #TOO MANY ROWS RIGHT?
  good_df = unique.data.frame(temp)
  
  
  ###MAKE A LIST OF COMBINED STATION DATA FOR EACH COUNTY
  mid_list = list()

  for (i in unique(reference_list$measuring_county)){
    
    df_new = data.frame(good_df[1,])
    subset = reference_list %>% filter(measuring_county == i) %>% select(county_code,site_number)
  
    for (j in 1:nrow(subset)){
      pull = good_df %>% filter(county_code == reference_list$county_code[j], site_number == reference_list$site_number[j])
      
      df_new = rbind(df_new,pull)
    }
    
    df_new = df_new[-1,]
    mid_list[[i]] = df_new
  }
  
  ###AGGREGATE DAILY DATA TO MONTHLY FOR EACH COUNTY
  
  final_list = lapply(mid_list,monthly_agg)
  
  return(final_list)
}
```


ASSEMBLING FINAL DATASETS

CERTAIN POLLUTANT STANDARDS APPEAR TO BE BETTER THAN OTHERS BC IT HAS VALUES FOR AQI 

Lead: Lead 3-Month 2009 ?? Has all NAs for AQI 

CO: CO 8-hour 1971

SO2: SO2 1-hour 2010

NO2: NO2 1-hour 2010

O3: Ozone 8-hour 2015 ; sample duration should be 8 HR 

PM10: PM10 24-hour 2006

PM2.5: PM25 24-hour 2012


```{r}
#Load raw data
pollutants1_2019 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Raw/pollutants1_2019_8.18.RData")
pollutants2_2019 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Raw/pollutants2_2019_8.18.RData")

Lead2019 = pollutants1_2019 %>% filter(parameter_code == "14129")
CO2019 = pollutants1_2019 %>% filter(parameter_code == "42101")
SO22019 = pollutants1_2019 %>% filter(parameter_code == "42401")
NO22019 = pollutants1_2019 %>% filter(parameter_code == "42602")
O32019 = pollutants2_2019 %>% filter(parameter_code == "44201")
PM102019 = pollutants2_2019 %>% filter(parameter_code == "81102")
PM2.52019 = pollutants2_2019 %>% filter(parameter_code == "88101")


Lead2019_final = raw_transform(rawdata = Lead2019,reference_list = Lead_stations[[6]],standard = "Lead 3-Month 2009")

CO2019_final = raw_transform(rawdata = CO2019,reference_list = CO_stations[[6]],standard = "CO 8-hour 1971")

SO22019_final = raw_transform(rawdata = SO22019,reference_list = SO2_stations[[6]],standard = "SO2 1-hour 2010")

NO22019_final = raw_transform(rawdata = NO22019,reference_list = NO2_stations[[6]],standard = "NO2 1-hour 2010")

O32019_final = raw_transform(rawdata = O32019,reference_list = O3_stations[[6]],standard = "Ozone 8-hour 2015")

PM102019_final = raw_transform(rawdata = PM102019,reference_list = PM10_stations[[6]],standard = "PM10 24-hour 2006") 

PM2.52019_final = raw_transform(rawdata = PM2.52019,reference_list = PM2.5_stations[[6]],standard = "PM25 24-hour 2012")
```




Note: In v4, Lead 2017 and PM2.5 2018 giving problematic aggregation... the station data quality check removed 1702 out of 2092 stations (for PM 2.5)

NO LONGER THE CASE IN THIS VERSION OF QUALITY CHECK

FILLING IN MISSING LEAD ENTRIES W 2016 and 2018 VALUES AVERAGED:

```{r}
#Which counties are lacking 
lacking_idx17 = c()

for (i in 1:58){
  if(nrow(Lead2017_final[[i]]) < 12){
    lacking_idx17 = c(lacking_idx17,i)
  }
}

#pull the data for those counties in 2017 and 2019 to fill in 2018
lacking16 = list()
lacking18 = list()
x = 1

for (j in lacking_idx17){
  lacking16[[x]] = Lead2016_final[[j]]
  lacking18[[x]] = Lead2018_final[[j]]
  x = x+1
}

#take average and fill it in 2018 dataset
for (k in 1:length(lacking16)){
  data1 = lacking16[[k]]
  data2 = lacking18[[k]]
  
  avg_data = data1
  avg_data$arithmetic_mean = (data1$arithmetic_mean + data2$arithmetic_mean)/2
  avg_data$aqi = (data1$aqi + data2$aqi)/2
  
  Lead2017_final[[lacking_idx17[k]]] = avg_data
}
```


FILLING IN MISSING PM 2.5 2018 ENTRIES W 2017 and 2019 VALUES AVERAGED: 

NOT NECESSARY FOR THIS VERSION OF QUALITY CHECK

```{r}
#Which counties are lacking 
lacking_idx18 = c()

for (i in 1:58){
  if(nrow(PM2.52018_final[[i]]) < 12){
    lacking_idx18 = c(lacking_idx18,i)
  }
}

#pull the data for those counties in 2017 and 2019 to fill in 2018
lacking17 = list()
lacking19 = list()
x = 1

for (j in lacking_idx18){
  lacking17[[x]] = PM2.52017_final[[j]]
  lacking19[[x]] = PM2.52019_final[[j]]
  x = x+1
}

#take average and fill it in 2018 dataset
for (k in 1:length(lacking17)){
  data1 = lacking17[[k]]
  data2 = lacking19[[k]]
  
  avg_data = data1
  avg_data$arithmetic_mean = (data1$arithmetic_mean + data2$arithmetic_mean)/2
  avg_data$aqi = (data1$aqi + data2$aqi)/2
  
  PM2.52018_final[[lacking_idx18[k]]] = avg_data
}
```



COMBINING EACH POLLUTANTS DATASET INTO A SINGLE DATAFRAME FOR THE YEAR 

```{r}
###Combine final data into one dataframe for 2014 
test69 = as.data.frame(do.call(rbind, Lead2019_final))
test69 = cbind(test69,rep(pollutants$parametercodes.code[1],nrow(test69))) #maybe change parameter codes to 1-7?
colnames(test69) = c("Year-Month","Value","AQI","Pollutant")

test70 = as.data.frame(do.call(rbind, CO2019_final))
test70 = cbind(test70,rep(pollutants$parametercodes.code[2],nrow(test70))) #maybe change parameter codes to 1-7?
colnames(test70) = c("Year-Month","Value","AQI","Pollutant")



# ###Just for Lead2017
# test69$`Year-Month` = test70$`Year-Month`



test71 = as.data.frame(do.call(rbind, SO22019_final))
test71 = cbind(test71,rep(pollutants$parametercodes.code[3],nrow(test71))) #maybe change parameter codes to 1-7?
colnames(test71) = c("Year-Month","Value","AQI","Pollutant")

test72 = as.data.frame(do.call(rbind, NO22019_final))
test72 = cbind(test72,rep(pollutants$parametercodes.code[4],nrow(test72))) #maybe change parameter codes to 1-7?
colnames(test72) = c("Year-Month","Value","AQI","Pollutant")

test73 = as.data.frame(do.call(rbind, O32019_final))
test73 = cbind(test73,rep(pollutants$parametercodes.code[5],nrow(test73))) #maybe change parameter codes to 1-7?
colnames(test73) = c("Year-Month","Value","AQI","Pollutant")

test74 = as.data.frame(do.call(rbind, PM102019_final))
test74 = cbind(test74,rep(pollutants$parametercodes.code[6],nrow(test74))) #maybe change parameter codes to 1-7?
colnames(test74) = c("Year-Month","Value","AQI","Pollutant")

test75 = as.data.frame(do.call(rbind, PM2.52019_final))
test75 = cbind(test75,rep(pollutants$parametercodes.code[7],nrow(test75))) #maybe change parameter codes to 1-7?
colnames(test75) = c("Year-Month","Value","AQI","Pollutant")



# ###Just for PM 2.5 2018
# test75$`Year-Month` = test74$`Year-Month`



#Combine each pollutant dataset into one big dataset for the year
final_data19 = rbind(test69,test70,test71,test72,test73,test74,test75)

###SAVE FINAL DATASET LOCALLY
saveRDS(final_data19,file = "C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Aggregated/final_data19_9.1.RData")
```

HAVE TO CLEAN DATA BEFORE FINALIZING: 

FIND MAX AQI (AMONG THE 7 POLLUTANTS) FOR EACH MONTH -> SET AS ACTUAL AQI FOR THAT MONTH

```{r}
#Do for each year 
months = c("01","02","03","04","05","06","07","08","09","10","11","12")

###Do for each year 
for (i in 1:58){
  idx1 = which(stringr::str_starts(rownames(final_data19), counties[i]))
  subset1 = final_data19[idx1,]
  subset1$`Year-Month`= as.Date(subset1$`Year-Month`)
  
  for (j in months){ 
  #Filter by county and date
    date = paste0("2019-",j,"-01")
    date = as.Date(date)
    subset2 = subset1 %>% filter(`Year-Month` == as.Date(date)) 
    
    trueAQI = max(na.omit(subset2$AQI))
    
    idx2 = which(subset1$`Year-Month` == date)
    subset1$AQI[idx2] = trueAQI
  }
  
  final_data19[idx1,] = subset1
}

###SAVE FINAL DATASET LOCALLY
saveRDS(final_data19,file = "C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Aggregated/final_data19_9.1.RData")
```




COMBINING EACH YEARS DATASET INTO ONE BIG TIDY DATAFRAME FOR AIR QUALITY COVARIATES 

```{r}
final_data14 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Aggregated/final_data14_9.1.RData")
final_data15 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Aggregated/final_data15_9.1.RData")
final_data16 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Aggregated/final_data16_9.1.RData")
final_data17 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Aggregated/final_data17_9.1.RData")
final_data18 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Aggregated/final_data18_9.1.RData")
final_data19 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Aggregated/final_data19_9.1.RData")

final_EPA_data = rbind(final_data14,final_data15,final_data16,final_data17,
                       final_data18,final_data19)
saveRDS(final_EPA_data,file = "C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Final/final_EPA_data_9.1.RData")


final_EPA_data = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Final/final_EPA_data_9.1.RData")
```


EXAMPLE OF CALCULATING ONE ENTRY IN K MATRIX 

```{r}
# A = test69 %>% filter(Time == "2014-01-01") %>% select(AQI)
# B = test69 %>% filter(Time == "2014-02-01") %>% select(AQI)
# 
# dist = mean((test70$AQI-test71$AQI)^2) 
# #once all pollutants are included (stacked together), test70 and test71 will be long vector (7*58 x 1)

# A = final_EPA_data %>% filter(`Year-Month` == "2014-01-01") 
# B = final_EPA_data %>% filter(`Year-Month` == "2014-02-01") 
# 
# AQIa = A$AQI[1:58]
# AQIb = B$AQI[1:58]
# 
# ABtest = c((A$Value-B$Value)^2,(AQIa-AQIb)^2) #58 counties * 8 pollutants 
# Kij = mean(ABtest)

K = matrix(0,nrow=72,ncol=72)
i = 1
j = 1
sigma = 5

for(t1 in unique(final_EPA_data$`Year-Month`)){
  for (t2 in unique(final_EPA_data$`Year-Month`)){
    A = final_EPA_data %>% filter(`Year-Month` == t1)
    B = final_EPA_data %>% filter(`Year-Month` == t2)
    AQIa = A$AQI[1:58]
    AQIb = B$AQI[1:58]
    
    ABtest = c((A$Value-B$Value)^2,(AQIa-AQIb)^2) #58 counties * 8 pollutants 
    K[i,j] = exp(-mean(ABtest) / (2*sigma^2))
    
    j = j+1
  }
  
  j = 1
  i = i+1
}
```
