---
title: "Important Plots"
author: "Jeffrey Wu"
date: "2023-08-22"
output: pdf_document
---

```{r,include=FALSE}
library(dplyr) 
library(tidyverse)
library(lubridate)
library(stringr)
library(zoo)
library(ggplot2)
library(urbnmapr)
library(devtools)
library(readxl)
library(spdep)
library(sp)
library(INLA)
library(HMMpa)
library(invgamma)
library(maps)
library(ggmap)

county_map <- map_data("county", region = "california")
```


## HEATMAPS FOR TOTAL MORTALITY FOR EACH YEAR: 

```{r,message=FALSE}
years = 2014:2019
for (i in years){
  totaldeaths = total.respmortality %>% group_by(County,Year) %>% summarise(Total_Deaths = sum(Total_Deaths)) %>% filter(Year == i)
  totaldeaths$County <- tolower(totaldeaths$County)

  #Plot onto a heatmap
  
  # Merge county map data with the dataset containing deaths
  county_map_totaldeaths <- merge(county_map, totaldeaths, by.x = "subregion", by.y = "County")
  
  # Plot the heatmap using ggplot2
  heatmap_plot <- ggplot(county_map_totaldeaths, aes(x = long, y = lat, group = group)) +
    geom_polygon(aes(fill = Total_Deaths), color = "black", size = 0.2) +  # Heatmap fill with     county lines
    coord_fixed() +
    labs(title = paste0("Total Respiratory Related Deaths in ", i),
         fill = "Deaths") +
    scale_fill_gradient(low = "blue", high = "red") +
    theme_minimal()
  
  
  # Display the heatmap plot
  print(heatmap_plot)  
}
```


HEATMAP FOR TOTAL MORTALITY PER 100,000 FOR 2014: 

```{r}
totaldeaths = rep(0,58)

for (i in 1:58){
  total = mortality3 %>% group_by(County_of_Death) %>% filter(Year_of_Death == 2014) %>% filter(County_of_Death == counties[i]) %>% summarise(Total_Deaths = sum(Total_Deaths))
  
  totaldeaths[i] = total$Total_Deaths
}

totaldeaths = data.frame(totaldeaths)
totaldeaths = cbind(totaldeaths,counties)
colnames(totaldeaths) = c("Total_Deaths","County")

#2010-2019 population data for CA 
USpops = read.csv("CA_census_pops1019.csv")
CApops = USpops %>% filter(STNAME == "California") %>% select(CTYNAME,POPESTIMATE2014)
CApops = CApops[-1,]

#Scale total deaths to become rate: per 100,000
totaldeaths$Total_Deaths = totaldeaths$Total_Deaths*100000/CApops$POPESTIMATE2014
totaldeaths$County <- tolower(totaldeaths$County)

#Plot onto a heatmap
# Merge county map data with the dataset containing deaths
county_map_totaldeaths <- merge(county_map, totaldeaths, by.x = "subregion", by.y = "County")

# Plot the heatmap using ggplot2
heatmap_plot <- ggplot(county_map_totaldeaths, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = Total_Deaths), color = "black", size = 0.2) +  # Heatmap fill with county lines
  coord_fixed() +
  labs(title = "Respiratory Related Deaths (per 100,000 people) in 2014",
       fill = "Deaths") +
  scale_fill_gradient(low = "yellow", high = "red",limits = c(0,130)) +
  theme_minimal()


# Display the heatmap plot
print(heatmap_plot)  
```


## HEATMAPS OF AQI LEVELS WITH STATION LOCATIONS MARKED

Get average AQI value for each county for Sept 2014: 

```{r}
Sept_avgs = c()

for (i in 1:58){ 
  #Filter by county
  county14 = final_data14 %>% filter(stringr::str_starts(rownames(final_data14), counties[i])) %>% filter(`Year-Month` == "2014-09-01") 
  
  SeptAQI = max(na.omit(county14$AQI))
  Sept_avgs[i] = SeptAQI
}
```

Avg AQI values for Sept 2014: 

```{r}
pollutants1_2014 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Raw/pollutants1_2014_8.17.RData")
pollutants2_2014 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Raw/pollutants2_2014_8.17.RData")

stationlats = c(unique(pollutants1_2014$latitude),unique(pollutants2_2014$latitude))
stationlongs = c(unique(pollutants1_2014$longitude),unique(pollutants2_2014$longitude))

station_points = data.frame(stationlats,stationlongs)

#Initializing map and station locations
ca_map <- map_data("county", region = "california")

#Match population dataset with ca_map
Sept_avgs = data.frame(Sept_avgs)
Sept_avgs = cbind(unique(ca_map$subregion),Sept_avgs)
colnames(Sept_avgs) = c("subregion","AQI")

merged_data <- merge(ca_map, Sept_avgs, by = "subregion", all.x = TRUE)

#Plot
gg_with_grouping <- ggplot() +
  geom_polygon(data = merged_data, aes(x = long, y = lat, group = group, fill = AQI), 
               color = "black") +
  coord_fixed(ratio = 1.3, xlim = c(-125, -112), ylim = c(30, 42)) +
  theme_void() +
  labs(title = "Heatmap of AQI levels with Station Locations for Sept 2014") +
  scale_fill_gradient(low = "yellow", high = "red")

# Add points
gg_with_grouping <- gg_with_grouping +
  geom_point(data = station_points, aes(x = stationlongs, y = stationlats), 
             color = "blue", size = 1)

print(gg_with_grouping)
```




## HEATMAP OF POPULATION WITH STATION LOCATIONS MARKED (2015)

```{r}
pollutants1_2015 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Raw/pollutants1_2015_8.17.RData")
pollutants2_2015 = readRDS("C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/EPA data/Raw/pollutants2_2015_8.17.RData")

stationlats = c(unique(pollutants1_2015$latitude),unique(pollutants2_2015$latitude))
stationlongs = c(unique(pollutants1_2015$longitude),unique(pollutants2_2015$longitude))

station_points = data.frame(stationlats,stationlongs)

#Initializing map and station locations
ca_map <- map_data("county", region = "california")

#Match population dataset with ca_map
#2010-2019 population data for CA 
USpops = read.csv("CA_census_pops1019.csv")
CApops = USpops %>% filter(STNAME == "California") %>% select(CTYNAME,POPESTIMATE2015)
CApops = CApops[-1,]

CApops$CTYNAME = unique(ca_map$subregion)
colnames(CApops) = c("subregion","pop")

merged_data <- merge(ca_map, CApops, by = "subregion", all.x = TRUE)

#Plot
gg_with_grouping <- ggplot() +
  geom_polygon(data = merged_data, aes(x = long, y = lat, group = group, fill = pop), 
               color = "black") +
  coord_fixed(ratio = 1.3, xlim = c(-125, -112), ylim = c(30, 42)) +
  theme_void() +
  labs(title = "Heatmap of County Populations with Station Locations for 2015") +
  scale_fill_gradient(low = "lightblue", high = "darkblue")

# Add points
gg_with_grouping <- gg_with_grouping +
  geom_point(data = station_points, aes(x = stationlongs, y = stationlats), 
             color = "red", size = 1.5)

print(gg_with_grouping)
```



## PLOT MAP OF CA WITH STATION LOCATIONS AND CLUSTERS

```{r}
#Initializing map and station locations
ca_map <- map_data("county", region = "california")

#Stations measuring SO2 for 2014
station_points = data.frame(
  longitude = all_pollutants_station_2014data[[3]]$longitude,
  latitude = all_pollutants_station_2014data[[3]]$latitude
)

#Attach cluster labels to each county
clusterlabels = data.frame(CA_data_cluster$NAME,clus5$groups)
names(clusterlabels) = c("subregion","Cluster")

o = order(clusterlabels$subregion)
clusterlabels = clusterlabels[o,]
clusterlabels$subregion = tolower(clusterlabels$subregion)
clusterlabels$Cluster = factor(clusterlabels$Cluster)

merged_data <- merge(ca_map, clusterlabels, by = "subregion", all.x = TRUE)

# Define distinct colors for each category
distinct_colors <- c("yellow", "blue", "green", "purple", "orange")

#Plot
gg_with_grouping <- ggplot() +
  geom_polygon(data = merged_data, aes(x = long, y = lat, group = group, fill = Cluster), 
               color = "black") +
  coord_fixed(ratio = 1.3, xlim = c(-125, -112), ylim = c(30, 42)) +
  theme_void() +
  labs(title = "Clustered Map of California with Station Locations for ") +
  scale_fill_manual(values = distinct_colors)

# Add points
gg_with_grouping <- gg_with_grouping +
  geom_point(data = station_points, aes(x = longitude, y = latitude), 
             color = "red", size = 1.5)

print(gg_with_grouping)
```



## ADJACENCY GRAPH OF CA WITH NODES PLOTTED AT POP-WEIGHTED CENTROID OF EACH CLUSTER 

Identify pop weighted centroids for each cluster 

```{r}
#2010-2019 population data for CA 
USpops = read.csv("CA_census_pops1019.csv")
CApops = USpops %>% filter(STNAME == "California") %>% select(CTYNAME,POPESTIMATE2015)
CApops = CApops[-1,]


#Attach cluster labels to each county
clusterlabels = data.frame(CA_data_cluster$NAMELSAD,CA_data_cluster$INTPTLAT,CA_data_cluster$INTPTLON,clus5$groups)
names(clusterlabels) = c("CTYNAME","Lat","Long","Cluster")

o = order(clusterlabels$CTYNAME)
clusterlabels = clusterlabels[o,]
clusterlabels$Cluster = factor(clusterlabels$Cluster)

merged_data <- merge(CApops, clusterlabels, by = "CTYNAME", all.x = TRUE)
merged_data$Lat = as.numeric(merged_data$Lat)
merged_data$Long = as.numeric(merged_data$Long)


# Calculate Population-Weighted Centroids
centroid_data <- merged_data %>%
  group_by(Cluster) %>%
  summarize(
    WeightedLat = weighted.mean(Lat, POPESTIMATE2015),
    WeightedLong = weighted.mean(Long, POPESTIMATE2015)
  )
```

Make the plot: 

```{r}
library(leaflet)

adj_matrix <- matrix(c(
    0, 0, 1, 1, 1,
    0, 0, 1, 1, 1,
    1, 1, 0, 1, 1,
    1, 1, 1, 0, 1,
    1, 1, 1, 1, 0
), nrow = 5, byrow = TRUE)

#Nodes are specific cities 

# node_coords <- matrix(c(
#     34.0522, -118.2437,   # Los Angeles
#     37.7749, -122.4194,   # San Francisco
#     32.7157, -117.1611,   # San Diego
#     38.5816, -121.4944,   # Sacramento
#     36.7783, -119.4179    # Fresno
# ), nrow = 5, byrow = TRUE)

#Nodes are population weighted centroids of each cluster 

node_coords <- matrix(c(
    37.27924, -121.5303,   # Cluster 1
    38.97130, -121.2390,   # Cluster 2
    38.86888, -122.9104,   # Cluster 3
    34.24761, -117.8457,   # Cluster 4
    38.11163, -121.8772    # Cluster 5
), nrow = 5, byrow = TRUE)


california_map <- leaflet() %>% addTiles() %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6)


for (i in 1:5) {
    california_map <- addMarkers(
        california_map,
        lng = node_coords[i, 2],
        lat = node_coords[i, 1],
        popup = paste("Node", i)
    )
}

for (i in 1:4) {
    for (j in (i + 1):5) {
        if (adj_matrix[i, j] == 1) {
            california_map <- addPolylines(
                california_map,
                lng = c(node_coords[i, 2], node_coords[j, 2]),
                lat = c(node_coords[i, 1], node_coords[j, 1]),
                color = "blue"
            )
        }
    }
}

california_map
```



