---
title: "Mortality EM Algorithm"
author: "Jeffrey Wu"
date: "2023-08-29"
output: pdf_document
---

```{r}
library(dplyr) 
library(tidyverse)
library(lubridate)
library(stringr)
library(zoo)
library(ggplot2)
library(urbnmapr)
library(devtools)
library(readxl)
library(spdep)
library(sp)
library(INLA)
library(HMMpa)
library(invgamma)
library(pscl)
```

Load raw data: 

```{r}
mortality = read_xlsx("RespiratoryMortality1423.xlsx")
mortality = filter(mortality,Cause_of_Death %in% c("Chronic lower respiratory diseases","Influenza and pneumonia"))
mortality = filter(mortality,Year_of_Death < 2020)

Population = rep(100000,nrow(mortality))
mortality = cbind(mortality,Population)
```


Attach population for each age group to mortality: 

```{r}
population_age = read_xlsx("Population Categories.xlsx")

population_age = population_age[-(1:5),2]
population_age$...2 = as.numeric(population_age$...2)

#split under 5 category into < 1 and 1-4 years old 
less1 = floor(population_age$...2[1]*0.2)
onefour = floor(population_age$...2[1]*0.8)
death_byage = population_age$...2[-1]
death_byage = c(less1,onefour,death_byage)
death_byage = death_byage[1:19]

death_byage2 = death_byage[1:2]
idx = seq(from = 3, to = 17, by = 2)
for (i in idx){
  death_byage2 = c(death_byage2,(death_byage[i]+death_byage[i+1]))
}
death_byage2 = c(death_byage2,death_byage[19])


age_groups = unique(mortality$Age)
death_byage2 = data.frame(cbind(age_groups,death_byage2))
colnames(death_byage2) = c("Age_Group","Population_by_Age")
death_byage2$Population_by_Age = as.numeric(death_byage2$Population_by_Age)

#2010-2019 population data for CA 
USpops = read.csv("CA_census_pops1019.csv")
CApops = USpops %>% filter(STNAME == "California") %>% select(CTYNAME,POPESTIMATE2019)
counties = countycodes$value_represented #from EPA data file

weights = CApops[(2:59),2]
weights = weights/CApops[1,2]

groups = unique(mortality$Age)
step1 = 1
step2 = 1

for (i in counties){
  for (j in groups){
    idx = which(mortality$Age == j & mortality$County_of_Death == i)
    mortality$Population[idx] = ceiling(death_byage2$Population[step1]*weights[step2])
    step1 = step1+1
  }  
  step1 = 1
  step2 = step2+1
}

mortality$logpop = log(mortality$Population)
```


```{r}
censored_idx = which(mortality$Total_Deaths == "<11")
censorTF = mortality$Total_Deaths == "<11"
mortality = cbind(mortality,censorTF)
```



FUNCTION FOR IMPUTING CENSORED VALUE GIVEN LAMBDA

```{r}
impute_small_values = function(lambda){
  x = 1:10
  p = dpois(x,lambda)
  
  value = sum(x*p)/sum(p)
  return(value)
}
```


NEED TO GET INITIAL GUESS FOR THETA: AVG DEATHS (PER 100K PEOPLE) PER MONTH FOR ONE COUNTY

```{r}
uncensored_mortality = mortality %>% filter(censorTF == FALSE) %>% select(Total_Deaths,Population)
uncensored_mortality$Total_Deaths = as.numeric(uncensored_mortality$Total_Deaths)

theta = mean(uncensored_mortality$Total_Deaths*100000/uncensored_mortality$Population)
```

By using all the data, I obtained a crude initial guess for $\theta$ of 1.09


INITIAL IMPUTATION: 

```{r}
mortality2 = mortality
mortality2$Total_Deaths[censored_idx] = 0.01
mortality2$Total_Deaths = as.numeric(mortality2$Total_Deaths)

for (i in censored_idx){
  lambda = theta*mortality2$Population[i] / 100000
  deaths = impute_small_values(lambda)
  
  mortality2$Total_Deaths[i] = floor(deaths)
}
```


INITIAL REGRESSION MODELS: 

```{r}
mortality2$Age = factor(mortality2$Age)
mortality2$Cause_of_Death = factor(mortality2$Cause_of_Death)
mortality2$Month = factor(mortality2$Month)

pois_reg = glm(Total_Deaths ~ Age + Cause_of_Death + Month + offset(logpop), family = "poisson", data = mortality2)
ZIP_reg = zeroinfl(Total_Deaths ~ Age + Cause_of_Death + Month + offset(logpop) | 1, data = mortality2, dist = "poisson", link = "logit")

vec0 = coef(pois_reg)
# vec0 = coef(ZIP_reg)


summary(pois_reg)
summary(ZIP_reg)
```


START EM ALGORITHM UNTIL IMPUTED VALUES CONVERGE 

REGULAR POISSON (SHOULD BE SUFFICIENT)

```{r}
mortality3 = mortality2
model = pois_reg

model_diff = 100
iter = 1
vec0 = coef(model)

while((model_diff > 0.01) & (iter < 10)){
  
  #impute data (should be between 1-10)
  fvs = fitted.values(model)
  
  for (i in censored_idx){
    deaths = impute_small_values(fvs[i])
    mortality3$Total_Deaths[i] = floor(deaths)
  }
  
  #fit model on "new" data
  model = glm(Total_Deaths ~ Age + Cause_of_Death + Month + offset(logpop), family = "poisson", data = mortality3)
  vec1 = coef(model)
  
  model_diff = max(abs(vec1 - vec0))
  iter = iter+1
  vec0 = vec1
}

final_pois_reg = model
summary(final_pois_reg)

logLik(final_pois_reg)
```




ZIP MODEL (PROB UNNECESSARY BC W WAS ESTIMATED AS LIKE 0.0005)

```{r}
mortality4 = mortality2
model = ZIP_reg

model_diff = 100
iter = 1
vec0 = coef(model)

while((model_diff > 0.01) & (iter < 10)){
  
  #impute data (should be between 1-10)
  fvs = fitted.values(model)
  
  for (i in censored_idx){
    deaths = impute_small_values(fvs[i])
    mortality4$Total_Deaths[i] = floor(deaths)
  }
  
  #fit model on "new" data
  model = zeroinfl(Total_Deaths ~ Age + Cause_of_Death + Month + offset(logpop) | 1, data = mortality4, dist = "poisson", link = "logit")
  vec1 = coef(model)
  
  model_diff = max(abs(vec1 - vec0))
  iter = iter+1
  vec0 = vec1
}

final_zip = model
summary(final_zip)

logLik(final_zip)
# exp(-5.7266) / (1+exp(-5.7266))

lldiff = logLik(final_pois_reg) - logLik(final_zip)
```





