---
title: "LengthScale_Bounds"
author: "Jeffrey Wu"
date: "2024-06-13"
output: html_document
---

Finding reasonable upper bound for each model's length scale parameters 

# Proposed model 2 (EPA kernel)

```{r}
kernel1 = function(EPA_data = final_EPA_agg_data,t1,t2,rho){
  A = EPA_data %>% filter(Time == t1)
  B = EPA_data %>% filter(Time == t2)
  AQIa = unique(A$AQI)
  AQIb = unique(B$AQI)
  
  ABtest = c((A$Value-B$Value)^2,(AQIa-AQIb)^2) #7 clusters * 8 measurements 
  
  correlation = exp(- (sum(ABtest)^2)
                 / (2*rho)) * exp(- (2*sin(sum(abs(ABtest))*pi/12)^2)
                 / (rho))
  
  return(correlation)
}

#Less than 0.05 correlation at 3 yr lag
corr3 = kernel1(t1=1,t2=36,rho=1:1000)
plot(corr3)

length(which(corr3 < 0.05))

kernel1(t1=1,t2=36,rho=70)

#Less than 0.05 correlation at 5 yr lag
corr5 = kernel1(t1=1,t2=60,rho=1:1000)
plot(corr5)

length(which(corr5 < 0.05))

kernel1(t1=1,t2=60,rho=203)
```

If memoryless after 3 years, rho upper bound is 70

If memoryless after 5 years, rho upper bound is 203

# Proposed model 1 (time kernel)

```{r}
kernel2 = function(t1=1,t2=60,rho){
  correlation = exp(- (abs(t1-t2)^2) / (2*rho)) * exp(- (2*sin(sum(abs(t1-t2))*pi/12)^2)
                     / (rho))
  return(correlation)
}

#Less than 0.05 correlation at 3 yr lag
corr3 = kernel2(1,36,1:1000)
plot(corr3)

length(which(corr3 < 0.05))

kernel2(1,36,204)

#Less than 0.05 correlation at 5 yr lag
corr5 = kernel2(1,60,1:1000)
plot(corr5)

length(which(corr5 < 0.05))

kernel2(1,60,581)
```

If memoryless after 3 years, rho upper bound is 204

If memoryless after 5 years, rho upper bound is 581

# Proposed model 3 (EPA kernel * time kernel)

```{r}
kernel3 = function(EPA_data = final_EPA_agg_data,t1,t2,rho){
  A = EPA_data %>% filter(Time == t1)
  B = EPA_data %>% filter(Time == t2)
  AQIa = unique(A$AQI)
  AQIb = unique(B$AQI)
  
  ABtest = c((A$Value-B$Value)^2,(AQIa-AQIb)^2) #7 clusters * 8 measurements 
  
  correlation1 = exp(- (sum(ABtest)^2)
                 / (2*rho)) * exp(- (2*sin(sum(abs(ABtest))*pi/12)^2)
                 / (rho))
  
  correlation2 = exp(- (abs(t1-t2)^2) / (2*rho)) * exp(- (2*sin(sum(abs(t1-t2))*pi/12)^2)
                     / (rho))
  
  return(correlation1*correlation2)
}

#Less than 0.05 correlation at 3 yr lag
corr3 = kernel3(t1=1,t2=36,rho=1:1000)
plot(corr3)

length(which(corr3 < 0.05))

kernel3(t1=1,t2=36,rho=274)

#Less than 0.05 correlation at 5 yr lag
corr5 = kernel3(t1=1,t2=60,rho=1:1000)
plot(corr5)

length(which(corr5 < 0.05))

kernel3(t1=1,t2=60,rho=784)
```

If memoryless after 3 years, rho upper bound is 274

If memoryless after 5 years, rho upper bound is 784


# Proposed model 4 (EPA kernel + time kernel)

```{r}
kernel4 = function(EPA_data = final_EPA_agg_data,t1,t2,rho){
  A = EPA_data %>% filter(Time == t1)
  B = EPA_data %>% filter(Time == t2)
  AQIa = unique(A$AQI)
  AQIb = unique(B$AQI)
  
  ABtest = c((A$Value-B$Value)^2,(AQIa-AQIb)^2) #7 clusters * 8 measurements 
  
  correlation1 = exp(- (sum(ABtest)^2)
                 / (2*rho)) * exp(- (2*sin(sum(abs(ABtest))*pi/12)^2)
                 / (rho))
  
  correlation2 = exp(- (abs(t1-t2)^2) / (2*rho)) * exp(- (2*sin(sum(abs(t1-t2))*pi/12)^2)
                     / (rho))
  
  return(correlation1+correlation2)
}

#Less than 0.05 correlation at 3 yr lag
corr3 = kernel4(t1=1,t2=36,rho=1:800)
plot(corr3)

length(which(corr3 < 0.05))

kernel4(t1=1,t2=36,rho=69)

#Less than 0.05 correlation at 5 yr lag
corr5 = kernel4(t1=1,t2=60,rho=1:800)
plot(corr5)

length(which(corr5 < 0.05))

kernel4(t1=1,t2=60,rho=203)
```

If memoryless after 3 years, rho upper bound is 69

If memoryless after 5 years, rho upper bound is 203

# Proposed model 5 (AR kernel + DL kernel + Interaction kernel)

```{r}
kernel5 = function(data = decomposed_cluster_data,t1,t2,rho){
  #Calculating gram matrix K_AR
  K_AR_periodic_cluster = list()
  
  for (c in 1:num_clus){
    
    #Grab S_random data for cluster c
    cluster_data = decomposed_cluster_data[[c]]
    S_random_clus = cluster_data$S_random
    
    #Create a list to contain covariance matrix for each pollutant (8)
    K_AR_periodic_list = c()
    
    time_span = nrow(S_random_clus)
    
    #Calculate a AR 1 covariance matrix for each pollutant and store in list
    for (i in 1:8){
      ts =  S_random_clus[,i]
    
      K_covariate_periodic = 0

      if (abs(t1-t2) <= 1){
        
        K_covariate_periodic = exp(- ((ts[t1] - ts[t2])^2) #Locally periodic kernel 
                     / (2*rho)) * exp(- (2*sin((abs(ts[t1] - ts[t2]))*pi/12)^2)
                     / (rho))
      }else{
        K_covariate_periodic = 0
      }
      
    K_AR_periodic_list = c(K_AR_periodic_list,K_covariate_periodic)
    
    K_AR_periodic = 0
    
    #Add each pollutant's covariance matrix to get AR 1 matrix for each cluster
    for(i in 1:length(K_AR_periodic_list)){
      K_AR_periodic = K_AR_periodic + ((1/8)*K_AR_periodic_list[[i]])
    }
    
    K_AR_periodic_cluster[[c]] = K_AR_periodic
    }
  }  
  
  K_AR_periodic = 0
  
  for(i in 1:num_clus){
    K_AR_periodic = K_AR_periodic + ((1/num_clus)*K_AR_periodic_cluster[[i]])
  }
  

  ###Calculating gram matrix K_DL
  K_DL_periodic_cluster = list()
  
  for (c in 1:num_clus){
    
    #Grab S_DL data for cluster c
    cluster_data = decomposed_cluster_data[[c]]
    S_DL_clus = cluster_data$S_DL
    
    #Create a list to store covariance matrix for each DL 
    K_DL_periodic_list = c()
    
    dl_lags = c(3,6,12)
    tracker = 1
    
    for (i in dl_lags){
      
      K_DL_periodic = 0
      
      #Calculate DL covariance matrix for specified lag   
      if ((abs(t1-t2) == 0) || (abs(t1-t2) == i)){
        
        K_DL_periodic = exp(- (sum(S_DL_clus[t1,] - S_DL_clus[t2,])^2)
                             / (2*rho)) * exp(- (2*sin(sum(abs(S_DL_clus[t1,] - S_DL_clus[t2,]))*pi/12)^2)
                             / (rho))
        
      } 
      else{
        K_DL_periodic = 0
      }
      
      K_DL_periodic_list = c(K_DL_periodic_list,K_DL_periodic)
    }
    
    #Combine the 3 DL covariance matrices together
    K_DL_periodic = 0
    
    for(i in 1:length(K_DL_periodic_list)){
      K_DL_periodic = K_DL_periodic + ((1/3)*K_DL_periodic_list[[i]])
    }
    
    #Store DL(3,6,12) covariance matrix for each cluster 
    K_DL_periodic_cluster[[c]] = K_DL_periodic
  }
  
  K_DL_periodic = 0
  
  for(i in 1:num_clus){
    K_DL_periodic = K_DL_periodic + ((1/num_clus)*K_DL_periodic_cluster[[i]])
  }
  
  
  ###Calculating gram matrix K_Interaction
  K_Interaction_periodic_cluster = list()
  
  for (c in 1:num_clus){
    
    #Grab interaction pair data for cluster c
    cluster_data = decomposed_cluster_data[[c]]
    W2_clus = cluster_data$W2

    K_interaction_periodic_list = c()
    
    column_names = colnames(W2_clus)
    time_span = nrow(W2_clus)
    
    #Create sequence of indices corresponding to comparisons for real time and one lag interaction effects
    lag0_idx = seq(2,3601,by=61)
    lag1_idx = seq(1,3600,by=61)
    
    #Calculate a kernel for each interaction pair 
    for (a in 1:length(column_names)){
      interaction =  W2_clus[,a]
      
      #First calculate these two interaction kernels separately 
      K_int0_periodic = 0
      K_int1_periodic = 0

      #Locally periodic kernels 
      K_int0_periodic = exp(- ((interaction[lag0_idx[t1]] - interaction[lag0_idx[t2]])^2)
                      / (2*rho)) * 
        exp(- (2*sin((abs(interaction[lag0_idx[t1]] - interaction[lag0_idx[t2]]))*pi/12)^2)
                      / (rho))
          
      K_int1_periodic = exp(- ((interaction[lag1_idx[t1]] - interaction[lag1_idx[t2]])^2)
                           / (2*rho)) * 
        exp(- (2*sin((abs(interaction[lag1_idx[t1]] - interaction[lag1_idx[t2]]))*pi/12)^2)
                      / (rho))

      
      #Combine real time and one lag interaction kernels together
      K_interaction_periodic = 0.5*K_int0_periodic + 0.5*K_int1_periodic
      K_interaction_periodic_list = c(K_interaction_periodic_list,
                                      K_interaction_periodic)
    }
    
    #Combine kernels for each interaction pair together
    K_interaction_periodic = 0
    
    for(i in 1:length(K_interaction_periodic_list)){
      K_interaction_periodic = K_interaction_periodic + ((1/length(K_interaction_periodic_list))*K_interaction_periodic_list[[i]])
    }

    #Store final interaction kernel (for all pairs) for each cluster 
    K_Interaction_periodic_cluster[[c]] = K_interaction_periodic
  }
  
  K_interaction_periodic = 0
  
  for(i in 1:num_clus){
    K_interaction_periodic = K_interaction_periodic + ((1/length(K_Interaction_periodic_cluster))*K_Interaction_periodic_cluster[[i]])
  }

correlation = (1/3)*(K_AR_periodic + K_DL_periodic + K_interaction_periodic)

return(correlation)
}
```

```{r}
#Less than 0.05 correlation at 3 yr lag
corr3 = c()
grid = seq(0.001,0.1,length.out=100)

for (i in grid){
  corr3 = c(corr3,kernel5(data = decomposed_cluster_data, t1 = 1,t2=36,rho=i))
}

plot(corr3)

grid[length(which(corr3 < 0.05))]

kernel5(t1=1,t2=36,rho=0.013)

#Less than 0.05 correlation at 5 yr lag
corr5 = c()
grid = seq(0.001,0.01,length.out=100)

for (i in grid){
  corr5 = c(corr5,kernel5(data = decomposed_cluster_data, t1 = 1,t2=60,rho=i))
}

plot(corr5)

grid[length(which(corr5 < 0.05))]

kernel5(t1=1,t2=60,rho=0.005818182)
```
