---
title: "Likelihood Ratio Test"
author: "Jeffrey Wu"
date: "2023-08-16"
output: pdf_document
---

FILTER TO ONE AGE GROUP AND ONE CAUSE OF DEATH: 75-84 yr old / CLRD

Convert $\lambda$ which is population weighted rate to $\theta$ 

Let $\lambda =$ death rate per 100,000 people and let $\theta = \frac{\lambda N}{100000}$ (one per county)

Write likelihood for one county each month then sum over all months

## Truncated Poisson:

$P(X=0) = \frac{\theta^0 e^{-\theta}}{0!} = e^{-\theta}$

$P(1 \leq X \leq 10) = P(X \in TC) = \sum_{x=1}^{10}\frac{\theta^x e^{-\theta}}{x!} = e^{-\theta} p(\theta)$

$P(X=x) = \frac{\theta^x e^{-\theta}}{x!} for x \geq 11$

Combine these together to get the likelihood for our truncated Poisson distribution

$l(\vec{x_c}) = \Pi_{y=1}^6 \Pi_{m=1}^{12} (e^{-\theta})^{I(x_{cym} = 0 )} (p(\theta) e^{-\theta})^{I( x_{cym} \in TC )} (\frac{\theta^{x_{cym}} e^{-\theta}}{x_{cym} !})^{I(x_{cym} \geq 11)}$

Calculate log likelihood:

$log l(\vec{x_c}) = \sum_{y=1}^6 \sum_{m=1}^{12} [-\theta_c + log p(\theta_c)*I( x_{cym} \in TC ) + (x_{cym} log \theta_c - log(x_{cym}!))*I( x_{cym} \geq 11 ) ]$

```{r}
#Defining log likelihood for truncated Poisson 
model1_ll = function(x,theta){
  v = 1:10
  ptheta = sum(theta^v / gamma(v+1))
  ll = 0
    
  for(i in 1:length(x)){
    value = -theta + log(ptheta)*(x[i] > 0 & x[i] <= 10)  + (x[i]*log(theta))*(x[i] >= 11) #no need for constant term log(x[i]!)
    ll = ll + value
  }
  
  return(ll)
}
```

## Truncated ZIP

$P(X=0) = w + (1-w) e^{-\theta}$

$P(1 \leq X \leq 10) = P(X \in TC) = (1-w) \sum_{x=1}^{10} \frac{\theta^x e^{-\theta}}{x!} = (1-w) e^{-\theta} p(\theta)$

$P(X=x) = (1-w) \frac{\theta^x e^{-\theta}}{x!} for x \geq 11$

Combine these together to get the likelihood for our truncated Poisson distribution

$l(\vec{x_c}) = \Pi_{y=1}^6 \Pi_{m=1}^{12} (w + (1-w) e^{-\theta})^{I( x_{cym} = 0 )} ((1-w) e^{-\theta} p(\theta))^{I( x_{cym} \in TC )} ((1-w) \frac{\theta^x e^{-\theta}}{x_{cym}!})^{I( x_{cym} \geq 11 )}$

Calculate log likelihood:

$log l(\vec{x}) = \sum_{y=1}^6 \sum_{m=1}^{12} [log(w + (1-w) e^{-\theta})*I( x_{cym} = 0 ) + log(log(1-w) + log(p(\theta_c)) - \theta)*I( x_{cym} \in TC ) + (log(1-w) + x_{cym} log \theta_c - log(x_{cym}!))*I( x_{cym} \geq 11 ) ]$

```{r}
#Define log likelihood of truncated ZIP
model2_ll = function(x,theta,w){
  v = 1:10
  ptheta = sum( (theta^v) / (gamma(v+1)) )
  ll = 0
    
  for(i in 1:length(x)){
    
    value1 = (log(w + (1-w)*exp(-theta))) * (x[i] == 0) 
    value2 = (log(1-w) + log(ptheta) - theta) * (x[i] > 0 & x[i] <= 10)
    value3 = (log(1-w) + x[i]*log(theta) - theta) * (x[i] >= 11) #no need for constant term log(x[i]!)
  
    ll = ll + (value1 + value2 + value3)
  }
  
  return(ll)
}
```

## LRT function for Truncated Poisson vs ZIP

Load data:

```{r,include=FALSE,warning=FALSE}
library(dplyr) 
library(tidyverse)
library(lubridate)
library(stringr)
library(zoo)
library(ggplot2)
library(urbnmapr)
library(devtools)
library(readxl)
library(spdep)
library(sp)
library(INLA)
library(HMMpa)
library(invgamma)
```


```{r,cache=TRUE,include=FALSE}
mortality = read_xlsx("RespiratoryMortality1423.xlsx")
mortality = filter(mortality,Cause_of_Death %in% c("Chronic lower respiratory diseases","Influenza and pneumonia"))
mortality = filter(mortality,Year_of_Death < 2020)

colnames(mortality)[1:3] = c("Year","Month","County")

CA.counties2 = read.csv("counties.ca.data.csv")
CA_fips = CA.counties2[order(CA.counties2$county),]
ca.coordinates = data.frame(CA_fips$county, CA_fips$lat,CA_fips$lng, CA_fips$county_fips)
colnames(ca.coordinates) = c("County","Lat","Long","County_fips")
row.names(ca.coordinates) = NULL

#Attach county fips, lat and long coordinates to observations 
mortality2 <- left_join(ca.coordinates, mortality, by = "County")

idx0 = which(mortality2$Total_Deaths == "0")
idx11 = which(mortality2$Total_Deaths == "<11")

mortality2$Total_Deaths[idx11] = 3 #CHANGE BACK TO 0 AFTER LRT 

mortality2$Age = factor(mortality2$Age)
mortality2$Total_Deaths = as.numeric(mortality2$Total_Deaths)
```

Define this LRT into a function: 

Our hypotheses for these LRTs are: 

H0: standard Poisson model is appropriate fit for the mortality dataset

HA: a zero-inflated Poisson model is a better fit for the mortality dataset than the standard Poisson model


```{r}
ws = seq(0.01,0.99,length.out=100)

#agegroups are of the form: Less than 1 year , 55 - 64 years , 85 years and over
#cause can take values: Chronic lower respiratory diseases OR Influenza and pneumonia

PoissonLRT = function(dataset, agegroup = "55 - 64 years",cause = "Chronic lower respiratory diseases", county, alpha = 0.05){

  x = dataset %>% filter(Age == agegroup, Cause_of_Death == cause) %>% filter(County == county) %>% arrange(Year) 
  x = x$Total_Deaths
  maxval = max(x)
  if(maxval == 0){maxval = 1}

  thetas = seq(0.001,maxval,length.out=100) 
  
  ###Find maximum likelihood for model 1
  
  result1 = matrix(0,nrow=length(thetas),ncol=2)
  count = 0
  
  for (theta in thetas){
      count = count+1
      result1[count,] = c(theta, model1_ll(x,theta))
  }
  
  result1 = data.frame(result1)
  colnames(result1) = c("theta","ll1")
  idx1 = which(result1$ll1 == max(result1$ll1))
  
  ll1 = result1[idx1,]$ll1
  
  ###Find maximum likelihood for model 2
  
  result2 = matrix(0,nrow=length(thetas)*length(ws),ncol=3)
  count = 0
  
  for (theta in thetas){
    for (w in ws){
      count = count+1
      result2[count,] = c(theta, w , model2_ll(x,theta,w))
    }
  }
  
  result2 = data.frame(result2)
  colnames(result2) = c("theta","w","ll2")
  idx2 = which(result2$ll2 == max(result2$ll2))
  
  ll2 = result2[idx2,]$ll2
  
  ###Perform LRT 
  
  TS = 2*(ll2 - ll1) #distributed chi-sq df1
  pvalue = 1-pchisq(TS,df = 1)
  
  decision = (pvalue < alpha)
  
  result_vec = c(county,round(as.numeric(result1$theta[idx1]),2),
                 round(as.numeric(ll1),2),round(as.numeric(result2$theta[idx2]),2),
                 round(as.numeric(ll2),2),round(as.numeric(TS),2),
                 round(as.numeric(pvalue),2),decision) #TRUE means reject H0
  
  return(result_vec)
}

test = PoissonLRT(dataset = mortality2,county = "Alpine")
```


## PERFORM LRT FOR EVERY COUNTY

```{r}
LRT_results = matrix(NA,nrow = 58, ncol = 8)
counties = unique(mortality2$County)

for (i in 1:58){
  LRT_results[i,] = PoissonLRT(mortality2,agegroup = "75 - 84 years",county = counties[i])
}

LRT_results = data.frame(LRT_results)
colnames(LRT_results) = c("County","Theta Model 1","Likelihood Model 1",
                          "Theta Model 2","Likelihood Model 2","Test Statistic",
                          "p-value","Reject H0?")

###ROUND TO TWO DECIMAL PLACES
head(LRT_results,10)
```



##WHICHEVER MODEL WINS COMPARE WITH MODEL 3 (TRUNCATED ZIP MODEL FOR EACH QUARTER)

To maximize likelihood here, just call model2_ll 4x with the subsetted quarterly datasets and get 4 max lls -\> add those 4 max ll values together to get ll3
