---
title: "Main Mortality Analysis"
author: "Jeffrey Wu"
date: "2023-09-11"
output: pdf_document
---

DOWNLOAD CAL VIDA RESPIRATORY MORTALITY DATA

```{r}
library(dplyr) 
library(tidyverse)
library(lubridate)
library(stringr)
library(zoo)
library(ggplot2)
library(urbnmapr)
library(devtools)
library(readxl)
library(spdep)
library(sp)
library(INLA)
library(HMMpa)
library(invgamma)

mortality = read_xlsx("RespiratoryMortality1423.xlsx")
mortality = filter(mortality,Cause_of_Death %in% c("Chronic lower respiratory diseases","Influenza and pneumonia"))
mortality = filter(mortality,Year_of_Death < 2020)

colnames(mortality)[1:3] = c("Year","Month","County")

CA.counties2 = read.csv("counties.ca.data.csv")
CA_fips = CA.counties2[order(CA.counties2$county),]
ca.coordinates = data.frame(CA_fips$county, CA_fips$lat,CA_fips$lng, CA_fips$county_fips)
colnames(ca.coordinates) = c("County","Lat","Long","County_fips")
row.names(ca.coordinates) = NULL

#Attach county fips, lat and long coordinates to observations 
mortality2 <- left_join(ca.coordinates, mortality, by = "County")

# write.csv(mortality2,"C:/Users/jeffr/Desktop/Spatiotemporal + Causal Inference/Wildfire Paper 1 Code/respmortality1423.csv")
```

DATA EXPLORATION 

ORIGINAL EDA

```{r}
#Sample: how many 0s in Alameda for people < 1 yr old in 2015
respmortality %>% filter(County == "Alameda") %>% filter(Age == "Less than 1 year") %>% filter(Year == 2015) %>% select(Total_Deaths) %>% table()

#How many 0s each year for every counties (5 number summary)
props = c()
counties = unique(respmortality$County)
years = sort(unique(respmortality$Year))

for (i in counties){
  for (j in years){
      filter = respmortality %>% filter(County == i) %>% filter(Year == j) %>% select(Total_Deaths) %>% table()
      values = as.numeric(filter)
      prop = values[1]/sum(values) #proportion of 0s
      props = c(props,prop) 
  }
}

summary(props)
hist(props)

#Use to get proportion of 0s for different age groups over all counties all years  (and then different years)
filter = respmortality %>% filter(Age == "45 - 54 years") %>% select(Total_Deaths) %>% table()

# filter = respmortality %>% filter(Age == "45 - 54 years") %>% filter(Year == 2014) %>% select(Total_Deaths) %>% table()

print(filter)
#plot(filter)
values = as.numeric(filter)
values[1]/sum(values)
```

NEW EDA

```{r}
#Find proportion of cell response values >= 11 for each age group 
age_groups = unique(mortality2$Age)

for (i in age_groups){
  filter = mortality2 %>% filter(Year == 2014) %>% filter(Age == i) %>% select(Total_Deaths) %>% table()
  values = as.numeric(filter)
  prop = (values[1]+values[2])/sum(values)
  print(1-prop)
}

# Create heatmap of prop >= 11 for each county for each age group 
county_names = ca.coordinates$County
props11 = matrix(1,nrow = length(county_names),ncol = length(age_groups))

for (i in 1:length(county_names)){
  for (j in 1:length(age_groups)){
    filter = mortality2 %>% filter(County == county_names[i]) %>% filter(Age == age_groups[j]) %>% select(Total_Deaths) %>% table()
    values = as.numeric(filter)
    if (length(values) == 1){
      prop = values[1]/sum(values)
    } else{
      prop = (values[1]+values[2])/sum(values)
    }
    props11[i,j] = 1-prop
  }
}

props11 = data.frame(props11)
colnames(props11) = age_groups

#Plot some histograms (one for each county)
# for (i in county_names){
#   filter = mortality2 %>% filter(County == i) %>% select(Total_Deaths) %>% table()
#   plot(filter,type = "h",main = i,xlab = "Total respiratory deaths")
# }

# FOR AGE GROUP 55-64 
#Plot onto a heatmap
county_proportions <- data.frame(
  County = county_names,  # County names
  Proportion = props11$`55 - 64 years`,  # Proportions for each county
  stringsAsFactors = FALSE
)

county_proportions$County <- tolower(county_proportions$County)

# Load county map data for California
county_map <- map_data("county", region = "california")

# Merge county map data with the dataset containing proportions
county_map_proportions <- merge(county_map, county_proportions, by.x = "subregion", by.y = "County")

# Plot the heatmap using ggplot2
heatmap_plot <- ggplot(county_map_proportions, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = Proportion), color = "black", size = 0.2) +  # Heatmap fill with county lines
  coord_fixed() +
  labs(title = "Proportion of Deaths Each Month (2014-2019) Greater than 10 by County",
       fill = "Proportion") +
  scale_fill_gradient(low = "blue", high = "red") +
  theme_minimal()


# Display the heatmap plot
print(heatmap_plot)
```



REFORMATTING DATA 

```{r}
#Get rid of <11 response values 
### Get imputed dataset mortality3 from EM Algorithm file 
data = mortality3

data1 = data %>% filter(Cause_of_Death == "Chronic lower respiratory diseases")
data2 = data %>% filter(Cause_of_Death == "Influenza and pneumonia")

newdeaths = data1$Total_Deaths + data2$Total_Deaths
data1$Total_Deaths = newdeaths

respmortality = data1[,-5]
```


Aggregate data into total deaths and 55 or older

```{r}
agg.respmortality = respmortality[1,]
agg.respmortality$Age = as.character(agg.respmortality$Age)
rows2 = seq(1,nrow(respmortality),11)

#Creates total deaths by adding deaths of all age groups together 
for (i in rows2){
  agg.respmortality[i,] = respmortality[i,]
  agg.respmortality$Total_Deaths[i] = sum(respmortality$Total_Deaths[i:(i+10)])
  agg.respmortality$Age[i] = "Everyone"
}

agg.respmortality = na.omit(agg.respmortality)
rownames(agg.respmortality) = NULL
total.respmortality = agg.respmortality[,-c(6:9)]
```



```{r}
#Creates age group 55 or older 
agg.respmortality2 = respmortality[1,]
agg.respmortality2$Age = as.character(agg.respmortality2$Age)
rows3 = seq(8,43744,11)
x = 1

for (i in rows3){
  x = x+1
  agg.respmortality2 = rbind(agg.respmortality2,respmortality[i,])
  agg.respmortality2$Total_Deaths[x] = sum(respmortality$Total_Deaths[i:(i+3)])
  agg.respmortality2$Age[x] = "55 years or older"
}
agg.respmortality2 = agg.respmortality2[-1,]

agg.respmortality2 = na.omit(agg.respmortality2)
rownames(agg.respmortality2) = NULL
old.respmortality = agg.respmortality2[,-c(6:9)]

#How different are the two datasets? 
sum(old.respmortality$Total_Deaths == total.respmortality$Total_Deaths)
```




Reformat dataset into time series format (rows are counties, columns are months)

Total deaths: 

```{r}
months = unique(total.respmortality$Month_of_Death)
years = sort(unique(total.respmortality$Year_of_Death))
counties = unique(total.respmortality$County_of_Death)
x = 0

total.mortality.ts = matrix(1,nrow = 58, ncol = 72)

for (k in counties){
  county.ts = c()
  x = x+1
  
  for (i in years){
    for (j in months){
      deaths = total.respmortality %>% filter(County_of_Death == k) %>%  filter(Year_of_Death == i) %>% filter(Month_of_Death == j) %>% select(Total_Deaths) %>% as.numeric()
      county.ts = c(county.ts,deaths)
    }
  } 
  
  total.mortality.ts[x,] = county.ts
}

# How many counties have rows of all 0s? 33
# count = 0
# 
# for (i in 1:58){
#   row = na.omit(total.mortality.ts[i,])
# 
#   if (sum(row) == 0){
#     count = count+1
#   }
# }


#Label time series data 
total.mortality.ts = as.data.frame(total.mortality.ts)

dates = c()
x=1
for (i in years){
  for (j in months){
    dates[x] = sprintf("%1.0f/%1.0f",j,i)
    x = x+1
  }
}

colnames(total.mortality.ts) = dates
ID = c(1:58)
total.mortality.ts = cbind(ID,counties,total.mortality.ts)

#test
mean(na.omit(total.mortality.ts[,11]))

#Attach cluster designations from SoA Data Analysis
clusterlabels = data.frame(CA_data_cluster$NAME,clus5$groups)
names(clusterlabels) = c("counties","Cluster")

o = order(clusterlabels$counties)
clusterlabels = clusterlabels[o,]

total.mortality.ts = left_join(clusterlabels,total.mortality.ts,by = "counties")
```


HOW MANY 0s DOES EACH COUNTY HAVE?

```{r}
numzeros_total = c()
for (i in 1:58){
  numzeros_total[i] = length(which(total.mortality.ts[i,3:74] == 0))
}

numzeros_total
propzeros_total = numzeros_total/72

length(which(propzeros_total > 0.85))
countycodes$value_represented[which(propzeros_total > 0.75)]

hist(propzeros_total,breaks = 20 ,xlab = "Proportion of months with 0 deaths",main = "Do some counties have more strings of 0s than others?")
```


Old people deaths: 

```{r}
x = 0

old.mortality.ts = matrix(1,nrow = 58, ncol = 72)

for (k in counties){
  county.ts = c()
  x = x+1
  
  for (i in years){
    for (j in months){
      deaths = old.respmortality %>% filter(County_of_Death == k) %>%  filter(Year_of_Death == i) %>% filter(Month_of_Death == j) %>% select(Total_Deaths) %>% as.numeric()
      county.ts = c(county.ts,deaths)
    }
  } 
  
  old.mortality.ts[x,] = county.ts
}


# How many counties have rows of all 0s? 33
# count = 0
# 
# for (i in 1:58){
#   row = na.omit(old.mortality.ts[i,])
#   
#   if (sum(row) == 0){
#     count = count+1
#   }
# }

#Label time series data 
old.mortality.ts = as.data.frame(old.mortality.ts)

colnames(old.mortality.ts) = dates
old.mortality.ts = cbind(ID,counties,old.mortality.ts)

#test
mean(na.omit(old.mortality.ts[,11]))


#Attach cluster designations from SoA Data Analysis
old.mortality.ts = left_join(clusterlabels,old.mortality.ts,by = "counties")
```


HOW MANY 0s DOES EACH COUNTY HAVE?

```{r}
numzeros_old = c()
for (i in 1:58){
  numzeros_old[i] = length(which(old.mortality.ts[i,3:74] == 0))
}

numzeros_old
propzeros_old = numzeros_old/72

length(which(propzeros_old > 0.95))
countycodes$value_represented[which(propzeros_old > 0.75)]

hist(propzeros_old,breaks = 20 ,xlab = "Proportion of months with 0 deaths",main = "Do some counties have more strings of 0s than others?")
```


AGGREGATING (POP WEIGHTED MEAN) MORTALITY DATA INTO CLUSTERS AS OPPOSED TO EACH COUNTY

Total deaths: 

```{r}
#County populations by year pulled from SoA data
countypops = CA_data %>% filter(Year > 2013) %>% select(Total_Pop,County,Year) %>% unique()
countypops = cbind(countypops,Cluster = rep(clusterlabels$Cluster,each=6))

cluster_mortality_total = matrix(NA,nrow = 72,ncol = 5)

for (i in 1:5){
  cluster = total.mortality.ts %>% filter(Cluster == i)
  year = 2014
  
  for(j in 1:72){
    #Obtain a weighted mean based on population
    col = cluster[,j+3]
    
    pops = countypops %>% filter(Year == year,Cluster == i) %>% select(Total_Pop) 
    cluster.pop = sum(pops)
    cluster.popweights = pops/cluster.pop
    cluster_mortality_total[j,i] = weighted.mean(data.frame(col),cluster.popweights)
    
    if ((j>12) & (j<25)){
      year = 2015
    }
    
    else if ((j>24) & (j<37)){
      year = 2016
    }
    
    else if ((j>36) & (j<49)){
      year = 2017
    }
    
    else if ((j>48) & (j<61)){
      year = 2018
    }
    
    else if ((j>60) & (j<73)){
      year = 2019
    }
    
    else{
      year = 2014
    }
  }
}

colnames(cluster_mortality_total) = c("Cluster 1","Cluster 2",
                                "Cluster 3","Cluster 4",
                                "Cluster 5")
rownames(cluster_mortality_total) = colnames(total.mortality.ts[4:75])
cluster_mortality_total = data.frame(cluster_mortality_total)

# #Removes Sept-Dec 2014 NAs
# cluster_mortality_total = na.omit(cluster_mortality_total)
```



WHAT TYPE OF GRAPH DO WE GET IF WE RUN HUGE ON THE CLUSTER MORTALITY DATA? IS IT ALSO COMPLETELY DISCONNECTED LIKE FOR SOA DATA?

```{r}
out.glasso = huge(as.matrix(cluster_mortality_total),method="glasso")

glasso.ebic = huge.select(out.glasso,criterion = "ebic")
glasso.ebic$refit
```

Adjacency matrix is all 0s just like for SoA data if robust selection bootstrap is applied...
All 1s if no robust selection 


Old people deaths: 

```{r}
#County populations by year pulled from SoA data
countypops = CA_data %>% filter(Year > 2013) %>% select(Total_Pop,County,Year) %>% unique()
countypops = cbind(countypops,Cluster = rep(clusterlabels$Cluster,each=6))

cluster_mortality_old = matrix(NA,nrow = 72,ncol = 5)

for (i in 1:5){
  cluster = old.mortality.ts %>% filter(Cluster == i)
  year = 2014
  
  for(j in 1:72){
    #Obtain a weighted mean based on population
    col = cluster[,j+3]
    
    pops = countypops %>% filter(Year == year,Cluster == i) %>% select(Total_Pop) 
    cluster.pop = sum(pops)
    cluster.popweights = pops/cluster.pop
    cluster_mortality_old[j,i] = weighted.mean(data.frame(col),cluster.popweights)
    
    if ((j>12) & (j<25)){
      year = 2015
    }
    
    else if ((j>24) & (j<37)){
      year = 2016
    }
    
    else if ((j>36) & (j<49)){
      year = 2017
    }
    
    else if ((j>48) & (j<61)){
      year = 2018
    }
    
    else if ((j>60) & (j<73)){
      year = 2019
    }
    
    else{
      year = 2014
    }
  }
}

colnames(cluster_mortality_old) = c("Cluster 1","Cluster 2",
                                "Cluster 3","Cluster 4",
                                "Cluster 5")
rownames(cluster_mortality_old) = colnames(old.mortality.ts[4:75])
cluster_mortality_old = data.frame(cluster_mortality_old)

# #Removes Sept-Dec 2014 NAs
# cluster_mortality_old = na.omit(cluster_mortality_old)
```

MAKE A TIME SERIES FOR EACH CLUSTER: 

```{r}
plot(ts(cluster_mortality_total$Cluster.1),xlab = "2014-2019 (Months)",ylab = "Respiratory related deaths", main = "Monthly population weighted means for Cluster 1")

plot(ts(cluster_mortality_total$Cluster.2),xlab = "2014-2019 (Months)",ylab = "Respiratory related deaths", main = "Monthly population weighted means for Cluster 2")

plot(ts(cluster_mortality_total$Cluster.3),xlab = "2014-2019 (Months)",ylab = "Respiratory related deaths", main = "Monthly population weighted means for Cluster 3")

plot(ts(cluster_mortality_total$Cluster.4),xlab = "2014-2019 (Months)",ylab = "Respiratory related deaths", main = "Monthly population weighted means for Cluster 4")

plot(ts(cluster_mortality_total$Cluster.5),xlab = "2014-2019 (Months)",ylab = "Respiratory related deaths", main = "Monthly population weighted means for Cluster 5")
```




